### 2020-09-29 analysis/02_tables.Rmd

---
title: "Tables"
author: "Suparna Chaudhry, Marc Dotson, and Andrew Heiss"
date: "Last run: `r format(Sys.time(), '%F')`"
output: 
  html_document:
    code_folding: hide
    pandoc_args:
      - "--default-image-extension=png"
editor_options: 
  chunk_output_type: console
---

```{r packages, warning=FALSE, message=FALSE, include=FALSE}
library(tidyverse)
library(pander)
library(magrittr)
library(glue)
library(scales)
library(janitor)
library(here)
```

```{r load-data}
sim_clean <- read_rds(here("data", "derived_data", "sim_final.rds"))
results <- read_rds(here("data", "raw_data", "final_data.rds"))

org_attributes <- tribble(
  ~`Issue area`, ~`Relationship with government`, ~`Funding`, ~`Funding sources`, ~`Organizational practices`, ~`Organization`,
  "Emergency response", "Friendly", "Small private donors", "Small private donors", "Financial transparency", "Amnesty International",
  "Environment", "Criticized", "Government grants", "Wealthy donors", "Accountability", "Greenpeace",
  "Human rights", "Crackdown", "", "Government grants", "", "Oxfam",
  "Refugee relief", "", "", "", "", "Red Cross"
)

persona_attributes <- tribble(
  ~`Demographics`, ~`Politics and public affairs`, ~`Social views`,
  "Higher income (> US median ($61,372)), high school graduate, frequent religious attendance", "Liberal (1), follows national and international news often, has traveled internationally", "High social trust: Trusts political institutions, trusts charities, thinks people should be more charitable, frequently volunteers, donates once a month, has a history of personal activism, is a member of an association",
  "Lower income (< US median), high school graduate, frequent religious attendance", "Conservative (7), follows news, has traveled", "Low social trust: Does not trust political institutions or charities, thinks people should be less charitable, does not volunteer or donate often, has no history of personal activism, is not a member of an association",
  "Higher income, college graduate, frequent religious attendance", "Liberal, does not follow news, has not traveled", "",
  "Lower income, college graduate, frequent religious attendance", "Conservative, does not follow news, has not traveled", "",
  "Higher income, high school graduate, rare religious attendance", "", "",
  "Lower income, high school graduate, rare religious attendance", "", "",
  "Higher income, college graduate, rare religious attendance", "", "",
  "Lower income, college graduate, rare religious attendance", "", ""
)
```

# Organization attributes

## All possible conjoint attributes

```{r results="asis"}
org_attributes %>% 
  select(Organization, `Issue area`, `Organizational practices`, `Funding sources`, `Relationship with government`) %>% 
  pandoc.table.return(caption = 'Organization attributes varied in the experiment {#tbl:organization-attributes-full}',
                      justify = "lllll", split.tables = Inf) %T>%
  cat() %>%
  cat(file = here("analysis", "output", "tables", "tbl-organization-attributes-full.md"))
```

## Attributes varied in simulation

```{r results="asis"}
org_attributes %>% 
  select(`Issue area`, `Relationship with government`, `Funding`) %>% 
  pandoc.table.return(caption = 'Organization attributes varied in the simulation, resulting in 24 hypothetical organizations {#tbl:organization-attributes}',
                      justify = "lll") %T>%
  cat() %>%
  cat(file = here("analysis", "output", "tables", "tbl-organization-attributes.md"))
```

# Persona attributes

## Attributes varied in simulation

```{r results="asis"}
persona_attributes %>% 
  pandoc.table.return(caption = 'Individual attributes varied in the simulation, resulting in 64 persona profiles {#tbl:persona-attributes}',
                      justify = "lll", split.tables = Inf) %T>%
  cat() %>%
  cat(file = here("analysis", "output", "tables", "tbl-persona-attributes.md"))
```

# Example simulation output

```{r results="asis"}
example_personas <- c("persona2", "persona63")

example_persona_details <- sim_clean %>%
  filter(persona_id %in% example_personas) %>% 
  select(starts_with("persona")) %>% 
  slice(1:2)

example_persona_details %>% 
  select(-persona_id) %>% 
  pivot_longer(cols = !persona) %>% 
  pivot_wider(names_from = "persona", values_from = "value") %>% 
  select(-name) %>% 
  pandoc.table(justify = "ll")
```

```{r results="asis"}
example_persona_results <- sim_clean %>% 
  filter(persona_id %in% example_personas) %>% 
  mutate(org_funding = str_to_sentence(str_remove(org_funding, "Mostly funded by "))) %>% 
  mutate(org_clean = glue("{organization}: {org_issue}, {org_funding}, {org_relationship}")) %>% 
  mutate(persona_desc = recode(
    persona_id,
    "persona2" = "Lower income high school graduate who rarely attends religious services; liberal who reads and travels; doesn't trust or donate",
    "persona63" = "Higher income college graduate who attends religious services; conservative who doesn't read or travel; trusts and donates")
  ) %>%
  mutate(persona_clean = glue("{persona}: {persona_desc}")) %>% 
  select(persona_clean, share, org_clean) %>% 
  pivot_wider(names_from = "persona_clean", values_from = "share") %>% 
  adorn_totals(where = "row", name = "**Total**")

example_persona_results_small <- example_persona_results %>% 
  slice(c(1, 2, 3, 7, 8, 9, 16, 17, 25)) %>% 
  mutate(across(where(is.numeric), ~ percent_format(accuracy = 0.1)(.))) %>% 
  add_row(org_clean = "…", .after = 3) %>% 
  add_row(org_clean = "…", .after = 7) %>% 
  add_row(org_clean = "…", .after = 10) %>% 
  mutate(across(everything(), ~replace_na(., "…"))) %>% 
  rename(Organization = org_clean)

example_persona_results_small %>% 
  pandoc.table.return(caption = 'Sample simulation output {#tbl:sim-output}',
                      justify = "lcc", split.tables = Inf) %T>%
  cat() %>%
  cat(file = here("analysis", "output", "tables", "tbl-sim-output.md"))
```

# Sample details

```{r}
vars_to_summarize <- tribble(
  ~category, ~variable, ~clean_name,
  "Demographics", "Q5.12", "Gender",
  "Demographics", "Q5.17", "Age",
  "Demographics", "Q5.13", "Marital status",
  "Demographics", "Q5.14", "Education",
  "Demographics", "Q5.15", "Income",
  "Attitudes toward charity", "Q2.5", "Frequency of donating to charity",
  "Attitudes toward charity", "Q2.6", "Amount of donations to charity last year",
  "Attitudes toward charity", "Q2.7_f", "Importance of trusting charities",
  "Attitudes toward charity", "Q2.8_f", "Level of trust in charities",
  "Attitudes toward charity", "Q2.10", "Frequency of volunteering",
  "Politics, ideology, and religion", "Q2.1", "Frequency of following national news",
  "Politics, ideology, and religion", "Q5.7", "Traveled to a developing country",
  "Politics, ideology, and religion", "Q5.1", "Voted in last election",
  "Politics, ideology, and religion", "Q5.6_f", "Trust in political institutions and the state",
  "Politics, ideology, and religion", "Q5.2_f", "Political ideology",
  "Politics, ideology, and religion", "Q5.4", "Involvement in activist causes",
  "Politics, ideology, and religion", "Q5.8", "Frequency of attending religious services",
  "Politics, ideology, and religion", "Q5.9", "Importance of religion"
)

summarize_factor <- function(x) {
  output <- table(x) %>% 
    as_tibble() %>% 
    magrittr::set_colnames(., c("level", "count")) %>% 
    mutate(level = factor(level, levels = levels(x))) %>%
    mutate(prop = count / sum(count),
           nice_prop = scales::percent(prop))

  return(list(output))
}

participant_summary <- results %>% 
  select(one_of(vars_to_summarize$variable)) %>% 
  summarize_all(summarize_factor) %>% 
  pivot_longer(cols = everything(), names_to = "variable", values_to = "details") %>% 
  left_join(vars_to_summarize, by = "variable") %>% 
  unnest(details) %>% 
  mutate(level = as.character(level)) %>% 
  mutate(level = case_when(
    variable == "Q2.7_f" & level == "1" ~ "1 (not important)",
    variable == "Q2.7_f" & level == "7" ~ "7 (important)",
    variable == "Q2.8_f" & level == "1" ~ "1 (no trust)",
    variable == "Q2.8_f" & level == "7" ~ "7 (complete trust)",
    variable == "Q5.6_f" & level == "1" ~ "1 (no trust)",
    variable == "Q5.6_f" & level == "7" ~ "7 (complete trust)",
    variable == "Q5.2_f" & level == "1" ~ "1 (extremely liberal)",
    variable == "Q5.2_f" & level == "7" ~ "7 (extremely conservative)",
    variable == "Q5.15" & level == "Less than median" ~ "Less than 2017 national median ($61,372)",
    variable == "Q5.17" & level == "Less than median" ~ "Less than 2017 national median (36)",
    TRUE ~ level
  )) %>% 
  mutate(clean_name_shrunk = ifelse(clean_name == lag(clean_name), "", clean_name),
         clean_name_shrunk = ifelse(is.na(clean_name_shrunk), 
                                    clean_name[1], 
                                    clean_name_shrunk),
         category_shrunk = ifelse(category == lag(category), "", category),
         category_shrunk = ifelse(is.na(category_shrunk), 
                                    category[1], 
                                    category_shrunk))
```

```{r results="asis"}
participant_summary %>% 
  select("Category" = category_shrunk, "Question" = clean_name_shrunk, 
         "Response" = level, "N" = count, "%" = nice_prop) %>% 
  pandoc.table.return(caption = 'Summary of individual respondent characteristics {#tbl:sample-details}',
                      justify = "lllcc", split.tables = Inf) %T>%
  cat() %>%
  cat(file = here("analysis", "output", "tables", "tbl-sample-details.md"))
```

\

# Original computing environment

<button data-toggle="collapse" data-target="#sessioninfo" class="btn btn-primary btn-md btn-info">Here's what we used the last time we built this page</button>

<div id="sessioninfo" class="collapse">

```{r show-session-info, echo=TRUE, width=100}
devtools::session_info()

writeLines(readLines(file.path(Sys.getenv("HOME"), ".R/Makevars")))
```

</div>