### 2020-10-01 analysis/experiment.Rmd
Add experiment text

---
title: "Experiment"
author: "Suparna Chaudhry, Marc Dotson, and Andrew Heiss"
date: "Last run: `r format(Sys.time(), '%F')`"
output: 
  html_document:
    code_folding: hide
    pandoc_args:
      - "--default-image-extension=png"
editor_options: 
  chunk_output_type: console
---

# Description

This survey was hosted on [Qualtrics](https://www.qualtrics.com) and participants were recruited through [Centiment](https://www.centiment.co/).

# Survey experiment {.survey}

```{r child = "text/experiment.md"}
```