### 2021-04-17 analysis/02_tables.Rmd

---
title: "Tables"
author: "Suparna Chaudhry, Marc Dotson, and Andrew Heiss"
date: "Last run: `r format(Sys.time(), '%F')`"
output: 
  html_document:
    code_folding: hide
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.retina = 3,
                      tidy.opts = list(width.cutoff = 120),  # For code
                      options(width = 90),  # For output
                      fig.asp = 0.618, fig.width = 7, 
                      fig.align = "center", out.width = "85%")

options(dplyr.summarise.inform = FALSE,
        knitr.kable.NA = "")
```

```{r load-libraries-data, message=FALSE, warning=FALSE}
library(tidyverse)
library(targets)
library(kableExtra)
library(glue)
library(scales)
library(janitor)
library(here)

# Generated via random.org
set.seed(8316)

# Load data
# Need to use this withr thing because tar_read() and tar_load() need to see the
# _targets folder in the current directory, but this .Rmd file is in a subfolder
withr::with_dir(here::here(), {
  tar_load(personas)
  tar_load(orgs)
  tar_load(sim_final)
  tar_load(survey_results)
  tar_load(participant_summary)
})
```

# Organization attributes

## All possible conjoint attributes

```{r organization-attributes-full}
orgs$org_attributes %>% 
  select(Organization, `Issue area`, `Organizational practices`, `Funding sources`, `Relationship with government`) %>% 
  kbl(align = "lllll",
      caption = "Organization attributes varied in the experiment") %>% 
  kable_styling()
```

## Attributes varied in simulation

```{r organization-attributes}
orgs$org_attributes %>% 
  select(`Issue area`, `Relationship with government`, `Funding`) %>% 
  kbl(align = "lll",
      caption = "Organization attributes varied in the simulation, resulting in 24 hypothetical organizations") %>% 
  kable_styling()
```

# Persona attributes

## Attributes varied in simulation

```{r persona-attributes}
personas$persona_attributes %>% 
  kbl(align = "lll",
      caption = "Individual attributes varied in the simulation, resulting in 64 persona profiles") %>% 
  kable_styling()
```

# Example simulation output

```{r example-personas}
example_personas <- c("persona2", "persona63")

example_persona_details <- sim_final %>%
  filter(persona_id %in% example_personas) %>% 
  select(starts_with("persona")) %>% 
  slice(1:2)

example_persona_details %>% 
  select(-persona_id) %>% 
  pivot_longer(cols = !persona) %>% 
  pivot_wider(names_from = "persona", values_from = "value") %>% 
  select(-name) %>% 
  kbl(align = "ll",
      caption = "Example personas") %>% 
  kable_styling()
```

```{r sim-output}
example_persona_results <- sim_final %>% 
  filter(persona_id %in% example_personas) %>% 
  mutate(org_funding = str_to_sentence(str_remove(org_funding, "Mostly funded by "))) %>% 
  mutate(org_clean = glue("{organization}: {org_issue}, {org_funding}, {org_relationship}")) %>% 
  mutate(persona_desc = recode(
    persona_id,
    "persona2" = "Lower income high school graduate who rarely attends religious services; liberal who reads and travels; doesn't trust or donate",
    "persona63" = "Higher income college graduate who attends religious services; conservative who doesn't read or travel; trusts and donates")
  ) %>%
  mutate(persona_clean = glue("{persona}: {persona_desc}")) %>% 
  select(persona_clean, share, org_clean) %>% 
  pivot_wider(names_from = "persona_clean", values_from = "share") %>% 
  adorn_totals(where = "row", name = "Total")

example_persona_results_small <- example_persona_results %>% 
  slice(c(1, 2, 3, 7, 8, 9, 16, 17, 25)) %>% 
  mutate(across(where(is.numeric), ~ percent_format(accuracy = 0.1)(.))) %>% 
  add_row(org_clean = "…", .after = 3) %>% 
  add_row(org_clean = "…", .after = 7) %>% 
  add_row(org_clean = "…", .after = 10) %>% 
  mutate(across(everything(), ~replace_na(., "…"))) %>% 
  rename(Organization = org_clean)

example_persona_results_small %>% 
  mutate(Organization = text_spec(Organization, bold = Organization == "Total")) %>% 
  kbl(align = "lcc",
      caption = "Sample simulation output",
      escape = FALSE) %>% 
  kable_styling()
```

# Sample details

```{r sample-details}
participant_summary %>% 
  select(Question = clean_name,
         Response = level,
         N = count,
         `%` = nice_prop) %>% 
  kbl(align = "lllcc",
      caption = "Summary of individual respondent characteristics") %>% 
  pack_rows(index = table(fct_inorder(participant_summary$category))) %>%
  collapse_rows(columns = 1, valign = "top") %>% 
  kable_styling()
```

\

# Original computing environment

<button data-toggle="collapse" data-target="#sessioninfo" class="btn btn-primary btn-md btn-info">Here's what we used the last time we built this page</button>

<div id="sessioninfo" class="collapse">

```{r show-session-info, echo=TRUE, width=100}
devtools::session_info()

writeLines(readLines(file.path(Sys.getenv("HOME"), ".R/Makevars")))
```

</div>