### 2021-04-22 README.Rmd
Update readme with full instructions

---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)

Title <- "Who Cares About Crackdowns? Exploring the Role of Trust in Individual Philanthropy"
Authors <- "Suparna Chaudhry, Marc Dotson, and Andrew Heiss"
Year <- "2021"
```

# `r Title`

[Suparna Chaudhry](http://www.suparnachaudhry.com/) • Department of International Affairs • Lewis & Clark College  
[Marc Dotson](https://marriottschool.byu.edu/directory/details?id=50683) • Marriott School of Business • Brigham Young University  
[Andrew Heiss](https://www.andrewheiss.com/) • Andrew Young School of Policy Studies • Georgia State University

---

[![OSF DOI](https://img.shields.io/badge/OSF-10.17605%2FOSF.IO%2FSM5EW-blue)](https://doi.org/10.17605/OSF.IO/SM5EW) [![DOI](https://zenodo.org/badge/DOI/10.5281/zenodo.4064840.svg)](https://doi.org/10.5281/zenodo.4064840)

> `r Authors`. `r Year`. ["`r Title`,"](https://doi.org/10.17605/OSF.IO/SM5EW)

**All this project's materials are free and open:**

- [Download the data](#data)
- [See the analysis notebook website](https://stats.andrewheiss.com/who-cares-about-crackdowns/)

![Preregistered](img/preregistered_large_color.png) &emsp; ![Open data](img/data_large_color.png) &emsp; ![Open](img/materials_large_color.png)

---

## Abstract

The phenomenon of closing civic space has adversely impacted the operations of INGOs. It has not just resulted in net losses of income for INGOs, but also how and where funds are allocated. We argue that individual private donors can be important in sustaining the operations of INGOs working in repressive contexts. Individual donors do not use the same performance-based metrics as official aid donors. Rather, trust can be an important component of individual donor support for nonprofits working towards difficult goals. How does trust in charitable organizations influence individuals' preferences to donate, especially when these groups face crackdown? Using a simulated market for philanthropic donations towards INGOs, based on data from a nationally representative sample of individuals in the United States who regularly donate to charity, we find that trust in charitable organizations matters substantially in shaping donor preferences. Donor profiles with high levels of social trust are likely to donate to INGOs with friendly relationships with host governments. This support holds steady if INGOs face criticism or crackdown. In contrast, donor profiles with lower levels of social trust prefer to donate to organizations that do not face criticism or crackdown abroad. The global crackdown on NGOs may thus possibly sour NGOs' least trusting individual donors. Our findings have practical implications for INGOs raising funds from individuals amid closing civic space.

---

This repository contains the data and code for our paper. Our preprint is online here:

> `r Authors`. `r Year`. "`r Title`"". Online at <https://doi.org/10.17605/OSF.IO/SM5EW>

## How to download and replicate

You can either [download the compendium as a ZIP file](/archive/master.zip) or use GitHub to clone or fork the compendium repository (see the green "Clone or download" button at the top of the GitHub page).

We use the [**renv** package](https://rstudio.github.io/renv/articles/renv.html) to create a stable version-specific library of packages, and we use the [**targets** package](https://docs.ropensci.org/targets/) to manage all file dependencies and run the analysis. To reproduce the findings and re-run the analysis, do the following:

1. Download and install these two fonts:
    - [IBM Plex Sans](https://fonts.google.com/specimen/IBM+Plex+Sans)
    - [Lora](https://fonts.google.com/specimen/Lora)
2. [Install R](https://cloud.r-project.org/) (and preferably [RStudio](https://www.rstudio.com/products/rstudio/download/#download)).
3. Open `whocares.Rproj` to open an [RStudio Project](https://r4ds.had.co.nz/workflow-projects.html).
4. Make sure you have a working installation of LaTeX:
    - *Easy-and-recommended way*: Install the **tinytex** package by running `install.packages("tinytex")` in the R console, then run `tinytex::install_tinytex()`
    - *Easy-but-requires-huge-4+-GB-download way*: Download TeX Live ([macOS](http://www.tug.org/mactex/); [Windows](https://miktex.org/))
5. If it's not installed already, R *should* try to install the **renv** package when you open the RStudio Project for the first time. If you don't see a message about package installation, install it yourself by running `install.packages("renv")` in the R console.
6. Run `renv::restore()` in the R console to install all the required packages for this project.
7. Run `targets::tar_make()` in the R console to automatically download all data files, process the data, run the analysis, and compile the paper and appendix.

Running `targets::tar_make()` will create several helpful outputs:

1. All project data in `data/`
2. An analysis notebook website in `analysis/_site/index.html`
3. PDF, HTML, and Word versions of the manuscript in `manuscript/output/`

## Data

This project includes the following data files:

- [**`data/raw_data/final_data.rds`**](https://osf.io/n2hwm/): Original results from the Qualtrics survey. This is [hosted at OSF](https://osf.io/n2hwm/) because of its size. Running `targets::tar_make(survey_results_file)` will download the `.rds` file from OSF and place it in `data/raw_data`. The [code for cleaning and processing this data is part of a separate project, "Why Donors Donate"](https://github.com/andrewheiss/why-donors-donate).
- [**`data/derived_data/survey_results.csv`**](data/derived_data/survey_results.csv): CSV version of the survey data.
- [**`data/derived_data/survey_results.yaml`**](data/derived_data/survey_results.yaml): [YAML metadata](https://csvy.org/) describing the syntax of the survey data.
- [**`data/raw_data/posterior_draws/public_political_social_charity_demo.rds`**](https://osf.io/msaz8/): Gamma (Γ) coefficients from our multilevel Bayesian model. This is [hosted at OSF](https://osf.io/msaz8/) because of its size. Running `targets::tar_make(gamma_draws_file)` will download the `.rds` file from OSF and place it in `data/raw_data/posterior_draws`. The [code for running this model is part of a separate project, "Why Donors Donate"](https://github.com/andrewheiss/why-donors-donate).
- [**`data/raw_data/Market Simulator Version 01.xlsx`**](data/raw_data/Market Simulator Version 01.xlsx): An interactive Excel version of the market simulator to help demonstrate the intuition behind all the moving parts of the simulation.
- [**`data/raw_data/data-xextT.csv`**](data/raw_data/data-xextT.csv): Per capita charitable donations in the US (2000–2014), downloaded from the [Datawrapper widget "How much money Americans give away"](https://theconversation.com/fewer-americans-are-giving-money-to-charity-but-total-donations-are-at-record-levels-anyway-98291#xextT) at ["Fewer Americans are giving money to charity but total donations are at record levels anyway"](https://theconversation.com/fewer-americans-are-giving-money-to-charity-but-total-donations-are-at-record-levels-anyway-98291).
- [**`data/raw_data/data-FTjUv.csv`**](data/raw_data/data-FTjUv.csv): Aggregate charitable donations in the US (1977–2017), downloaded from the [Datawrapper widget "Four decades of American charitable giving"](https://theconversation.com/fewer-americans-are-giving-money-to-charity-but-total-donations-are-at-record-levels-anyway-98291#FTjUv) at ["Fewer Americans are giving money to charity but total donations are at record levels anyway"](https://theconversation.com/fewer-americans-are-giving-money-to-charity-but-total-donations-are-at-record-levels-anyway-98291).

## Licenses

**Text and figures:** All prose and images are licensed under Creative Commons ([CC-BY-4.0](http://creativecommons.org/licenses/by/4.0/)).

**Code:** All code is licensed under the [MIT License](LICENSE.md).

## Contributions

We welcome contributions from everyone. Before you get started, please see our [contributor guidelines](CONTRIBUTING.md). Please note that this project is released with a [Contributor Code of Conduct](CONDUCT.md). By participating in this project you agree to abide by its terms.